Computable Poincar\'e--Friedrichs constants for the $L^{p}$ de Rham complex over convex domains and domains with shellable triangulations

### 2024-10-07 main.tex
complete literature review for the time being

@@

Not much attention however seems to have been given to the study of computable constants in~\eqref{math:intro:Wcurl_div}, even if the domain $\Omega$ is convex. 
Important results in the Hilbert case $p=2$ can be found in~\cite{guerini2004eigenvalue, Paul_Vald_PF_grad_curl_div_20} that we discuss now in the broader setting of the exterior calculus formalism, 
but the Banach case $p \neq 2$ seems largely unexplored.

Still in the Hilbert space case $p=2$, the Friedrichs inequality~\cite{burenkov1998sobolev} over the Sobolev space
with homogeneous Dirichlet boundary conditions on $\partial \Omega$ implies~\eqref{math:intro:Wdiv} via duality. 
However, that easy observation seems restricted to $p=2$.

The two
Poincar\'e--Friedrich inequalities for vector fields have appeared under different names. 
Let us assume for a moment that $\Omega$ is a weakly Lipschitz domain with trivial topology, and consider only the Hilbert case $p=2$. 
Then the constant in \eqref{math:intro:Wcurl:pf} agrees with the constant in the so-called Poincar\'e--Friedrichs--Weber inequality
\begin{align}
    \label{math:intro:weber}
    \| \bfu\|_{L^{2}(\Omega)}   \leq C_{\curl,\Omega,p} \| \curl \bfu \|_{L^{2}(\Omega)},
\end{align}
valid for all $\bfu \in \bfW^{2}(\curl,\Omega) \cap \bfW^{2}(\divergence,\Omega)$ that satisfy $\divergence \bfu = 0$ and have vanishing normal or vanishing tangential trace along $\partial \Omega$. 
Equivalently, \eqref{math:intro:weber} is valid for all $\bfu \in \bfW^{2}(\curl,\Omega)$ that are $L^{2}$-orthogonal to the gradients of scalar fields in $W^{1,2}(\Omega)$
or that have vanishing tangential trace and are $L^{2}$-orthogonal to the gradients of those scalar fields in $W^{1,2}(\Omega)$ that satisfy Dirichlet boundary conditions.
% Yet another rewriting is 
% \label{math:PF_orth} \begin{align}
%     \| \bfu\|_{L^{2}(\Omega)} & \leq C_{\curl,\Omega,2} \| \curl \bfu \|_{L^{2}(\Omega)}, 
% 	\label{math:PF_orth_curl}
% 	\\
%     \| \bfu\|_{L^{2}(\Omega)} & \leq \tilde C_{\divergence,\Omega,2} \| \divergence \bfu \|_{L^{2}(\Omega)}, 
% 	\label{math:PF_orth_div}
% \end{align}
% where the former holds for all vector fields $\bfu \in \bfW^{2}(\curl,\Omega)$ such that $\int_{\Omega} \bfu(x) \cdot \bfv(x) \;dx = 0$ for all $\bfv \in \bfW^{2}(\curl,\Omega)$ with $\curl \bfv = 0$,
% and where the latter holds for all functions $\bfu \in \bfW^{2}(\divergence,\Omega)$ such that $\int_{\Omega} \bfu(x) \cdot \bfv(x) \;dx = 0$ for all $\bfv \in \bfW^{2}(\divergence,\Omega)$ with $\divergence \bfv = 0$. 
We refer to Equation~(5) in~\cite{Fried_diff_forms_55}, Equation~(2) in \cite{Gaff_Hilbert_harm_55}, \cite{Web_compact_Maxw_80} as well as \cite[Lemmas~3.4 and~3.6]{Gir_Rav_NS_86}, \cite[Proposition~7.4]{Fer_Gil_Maxw_BC_97}, and the references therein. 
The general case of $L^p$ differential forms over Lipschitz manifolds subject to partial boundary conditions is discussed in~\cite{goldshtein2011hodge}.

However, while many of the above results rely on non-constructive estimates, 
we are interested in practically computable upper bounds.

>>>>

The qualitative existence of Poincar\'e--Friedrichs inequalities for the differential operators of vector calculus and exterior calculus is known for a large class of domains. 
There is a particularly large body of literature on the Poincar\'e inequality for the gradient and its numerous variants,
which may include weighted integrals or boundary terms. 
These Poincar\'e-type inequalities are often derived via non-constructive arguments, such as the Rellich embedding theorem,
and thus the resulting constants are not explicitly computable.

% divergence 
The divergence operator has received less attention. 
In the Hilbert space case $p=2$, the Friedrichs inequality~\cite{burenkov1998sobolev} over the Sobolev space
with homogeneous Dirichlet boundary conditions along $\partial \Omega$ implies~\eqref{math:intro:Wdiv} via duality. 
However, that easy duality argument, which yields an explicit upper bound proportional to the domain diameter, seems inherently restricted to $p=2$.

% curl 
The core challenges can be found in the discussion of the curl operator in three dimensions,
where Poincar\'e--Friedrich inequalities have appeared under different names. 
Let us assume for a moment that $\Omega$ is a weakly Lipschitz domain with trivial topology, and consider only the Hilbert case $p=2$. 
Then the constant in \eqref{math:intro:Wcurl:pf} agrees with the constant in the so-called Poincar\'e--Friedrichs--Weber inequality
\begin{align}
    \label{math:intro:weber}
    \| \bfu\|_{L^{2}(\Omega)}   \leq C_{\curl,\Omega,p} \| \curl \bfu \|_{L^{2}(\Omega)},
\end{align}
valid for all $\bfu \in \bfW^{2}(\curl,\Omega) \cap \bfW^{2}(\divergence,\Omega)$ that satisfy $\divergence \bfu = 0$ and have vanishing normal or vanishing tangential trace along $\partial \Omega$. 
Equivalently, \eqref{math:intro:weber} is valid for all $\bfu \in \bfW^{2}(\curl,\Omega)$ that are $L^{2}$-orthogonal to the gradients of scalar fields in $W^{1,2}(\Omega)$
or that have vanishing tangential trace and are $L^{2}$-orthogonal to the gradients of those scalar fields in $W^{1,2}(\Omega)$ that satisfy Dirichlet boundary conditions.
% Yet another rewriting is 
% \label{math:PF_orth} \begin{align}
%     \| \bfu\|_{L^{2}(\Omega)} & \leq C_{\curl,\Omega,2} \| \curl \bfu \|_{L^{2}(\Omega)}, 
% 	\label{math:PF_orth_curl}
% 	\\
%     \| \bfu\|_{L^{2}(\Omega)} & \leq \tilde C_{\divergence,\Omega,2} \| \divergence \bfu \|_{L^{2}(\Omega)}, 
% 	\label{math:PF_orth_div}
% \end{align}
% where the former holds for all vector fields $\bfu \in \bfW^{2}(\curl,\Omega)$ such that $\int_{\Omega} \bfu(x) \cdot \bfv(x) \;dx = 0$ for all $\bfv \in \bfW^{2}(\curl,\Omega)$ with $\curl \bfv = 0$,
% and where the latter holds for all functions $\bfu \in \bfW^{2}(\divergence,\Omega)$ such that $\int_{\Omega} \bfu(x) \cdot \bfv(x) \;dx = 0$ for all $\bfv \in \bfW^{2}(\divergence,\Omega)$ with $\divergence \bfv = 0$. 
We refer to Equation~(5) in~\cite{Fried_diff_forms_55}, Equation~(2) in \cite{Gaff_Hilbert_harm_55}, \cite{Web_compact_Maxw_80} as well as \cite[Lemmas~3.4 and~3.6]{Gir_Rav_NS_86}, \cite[Proposition~7.4]{Fer_Gil_Maxw_BC_97}, and the references therein. 
The general case of $L^p$ differential forms over Lipschitz manifolds subject to partial boundary conditions is discussed in~\cite{goldshtein2011hodge}.
However, while many of the above results rely on non-constructive estimates, 
we are interested in practically computable upper bounds.

@@
